---
title: "Introduction to R"
output:
  pdf_document: default
  html_notebook: default
---

### An Introduction to R

**OBJECTIVES**

- Basic Data Manipulation with `tidyverse` verbs
- Visualization with `ggplot`

### DataFrames in `R`

Here we load and read a `.csv` file.  This will create a dataframe object similar to what we have in python. 

### `tidyverse`

To install packages we can use the console, and once we have the library installed we will use the `dplyr` verbs: `select()`, `filter()`, `mutate()`, `group_by()`, `summarize()`. Here is a handy [cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).
```{r}
2 + 3
```


```{r}
#load library
library(tidyverse)
```
```{r}
head(data_ready_for_modelling)
```
```{r}
#select -- column selection
head(data_ready_for_modelling)
select(data_ready_for_modelling, Daily_Sum, fracs_per_day)
```

```{r}
#filter -- filter rows on condition
filter(data_ready_for_modelling, fracs_per_day > 20)
```

```{r}
#filter then select
data_ready_for_modelling %>%
  filter(fracs_per_day > 20) %>%
  select(Daily_Sum, volume_per_day)
```

```
```{r}
#mutate
#notice, if we run a head(), this new sepal.ten column is not there. IT doesn't get saved
#unless we assign it to a new variable
new_data <- data_ready_for_modelling %>%
  mutate(volume_per_day_1000 = volume_per_day / 1000)
head(new_data)
```

```
```{r}
#mutate
#notice, if we run a head(), this new sepal.ten column is not there. IT doesn't get saved
#unless we assign it to a new variable
new_data <- data_ready_for_modelling %>%
  mutate(volume_per_day_1000 = volume_per_day / 1000)
head(new_data)
```

```{r}
#group_by and summarize by counting: n() means count
new_data %>%
  group_by(fracs_per_day) %>%
  summarize(n())
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

### Introducing `ggplot`

In the R notebooks we can insert code blocks including loading libraries and generating plots.  Below we load in the tidyverse library and examine the built in dataset `mpg`.


```{r}
library(tidyverse)
head(new_data)
```

### Bar Plot

```r
ggplot(data, aes(x = column)) + 
    geom_bar()
```

```{r}
#aes is for aesthetic
ggplot(new_data, aes(x = fracs_per_day)) +
  geom_bar()
```


### Histogram

```r
ggplot(data, aes(x = column)) + 
    geom_histogram()
```

```{r}
ggplot(new_data, aes(x = fracs_per_day)) +
  geom_histogram() + 
  labs(title = 'Fracs per Day')
```


### Density Plot

```r
ggplot(data, aes(x = column)) + 
    geom_histogram()
```

```{r}
ggplot(new_data, aes(x = X1)) + 
  geom_density()
```


### Boxplots and Jitter

```r
ggplot(data, aes(x = x, y = y)) + 
    geom_boxplot() + 
    geom_jitter()
```

```{r}
ggplot(mpg, aes(x = manufacturer, y = displ )) +
  geom_boxplot() +
  geom_jitter() +
  labs(title = 'Displacement by Manufacturer')
```


### Scatter Plot

```r
ggplot(data, aes(x = x, y = y)) + 
    geom_point() + 
    geom_smooth()
```

```{r}
ggplot(new_data, aes(x = X1, y = fracs_per_day)) +
  geom_point(aes(color = Daily_Avg)) +
  geom_smooth(method = 'lm')
```


### A Basic Model

```r
lm(y ~ x)
```

```{r}
#first name the target, y, then name the x
#this is the equation for the line represented in the plot above
lm(mpg$hwy ~ mpg$displ)
```

```{r}

```





